---
title: "Data Analysis Project: PUBPOL 599 Goverance Analytics"
author: "Melissa Greenaway"
date: "March 11th, 2017"
output: html_document
---
First, import the large data set (it will take a few minutes!). The Financial Inclusion Insights (FII) survey is conducted by Intermedia, Inc. in partnership with the Gates Foundation, and asks respondents in 6 countries a series of questions related to their living standards and their use of digital financial services (DFS), like mobile money. This will help researchers and the public understand the progress being made in developing countries towards financial inclusion. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#set working directory
setwd("R:/Project/EPAR/Working Files/RA Working Folders/Melissa G/R and Python")

# packages needed for data cleaning, formatting, and analysis
# install.packages("haven")
# install.packages("sjmisc")
# install.packages("survey")
# install.packages("maptools")
# install.packages("RColorBrewer")
# install.packages("classInt")
# install.packages("scales")
# install.packages("ggmap")
# install.packages("ggplot2")

library(haven)
folder="Data"
fileName="Cross_WaveMaster_MG.dta"
fileToRead=file.path(folder,fileName)
dataStata=read_dta(fileToRead)
```

## Trim data set

Now, identify the variables we might be interested in, and create a new data frame with those variables:

```{r Trim data set}
varsOfInterest=c("weight", "wave","dfs_adopt","urban_rural","employed","phone_own","ppi_score","bank_own","age","ppi_cutoff","female","ed_level","literate","numerate","registered_MM","sim_ownoraccess","married","dfs_aware_gen","n_household","country")
crosswavesub=as.data.frame(dataStata)[varsOfInterest]
head(crosswavesub)
```

## Save as R data file

Now we can save the file as an R-data structure, to speed up analysis.

```{r Save as R file}
save(crosswavesub, file="crosswavesub.RData")
load("crosswavesub.RData")
head(crosswavesub)
```
## Data Structure
In viewing the data's structure, we see that each variable is "Atomic", and contains STATA metadata.
```{r structure of stata data}
str(crosswavesub)

```
# Data Cleaning and Formatting
To begin cleaning, we'll first change all NaNs to "NA"
```{r removing NAs}
crosswavesub[crosswavesub=='NaN']=NA # this is an actual change to the data frame
```

Now we can format the variables:
```{r Data formatting, message=FALSE, results='hide'}
# Data Formatting package
library(sjmisc)

#country categories
labelvar=c(get_labels(crosswavesub$country)) # can save factor labels
labelvar
capture.output(to_factor(crosswavesub$country))
#fix country
crosswavesub$country=to_label(crosswavesub$country) # this is the command that works
```

I'm going to change each variable to a factor (they're atomic right now)

```{r Data formatting 2}
library(sjmisc)
levels(crosswavesub$country)
str(crosswavesub$country)
#urban rural - These are binary categories
crosswavesub$urban_rural=to_factor(crosswavesub$urban_rural)
#employed
crosswavesub$employed=to_factor(crosswavesub$employed)
#phone_own
crosswavesub$phone_own=to_factor(crosswavesub$phone_own)
#bank_own
crosswavesub$bank_own=to_factor(crosswavesub$bank_own)
#ppi_cutoff
crosswavesub$ppi_cutoff=to_factor(crosswavesub$ppi_cutoff)
#married
crosswavesub$married=to_label(crosswavesub$married)
#literate
crosswavesub$literate=to_factor(crosswavesub$literate)
#registered MM
crosswavesub$registered_MM=to_factor(crosswavesub$registered_MM)
#sim own or access
crosswavesub$sim_ownoraccess=to_factor(crosswavesub$sim_ownoraccess)
#numerate
crosswavesub$numerate=to_factor(crosswavesub$numerate)
#dfs adopt
crosswavesub$dfs_adopt=to_factor(crosswavesub$dfs_adopt) # this is the one
#wave
crosswavesub$wave=to_factor(crosswavesub$wave)
#age
is.numeric(crosswavesub$age) # making sure it's numeric
#n_household
is.numeric(crosswavesub$n_household) # same as above
# female
crosswavesub$female=to_factor(crosswavesub$female)
# Changing ed_level to factor, ordered
crosswavesub$ed_level=factor(crosswavesub$ed_level,ordered = T)
str(crosswavesub)
# changing 
```

## Subset Data and Set Survey Weights

Create subset of data frames with each wave of data (for mapping of DFS adoption/use) and a subset with each country in wave 3. We also need to set survey weights for each subset of the data.
```{r Subset/Survey Weights }
library(survey)

#subset wave one data
wave1data=subset(crosswavesub,crosswavesub$wave=="1")
head(wave1data)
w1weight= svydesign(id=~1, weights=wave1data$weight,data=wave1data)

#subset wave two data
wave2data=subset(crosswavesub,crosswavesub$wave=="2")
head(wave2data)
w2weight= svydesign(id=~1, weights=wave2data$weight,data=wave2data)

#subset wave three data
wave3data<-subset(crosswavesub,crosswavesub$wave=="3")
head(wave3data)
w3weight= svydesign(id=~1, weights=wave3data$weight,data=wave3data)

#subset wave three data by country, for easier analysis
kenya3data<-subset(wave3data,wave3data$country=="Kenya")
k3weight= svydesign(id=~1, weights=kenya3data$weight,data=kenya3data)

nig3data<-subset(wave3data,wave3data$country=="Nigeria")
n3weight= svydesign(id=~1, weights=nig3data$weight,data=nig3data)

ind3data<-subset(wave3data,wave3data$country=="India")
ind3weight= svydesign(id=~1, weights=ind3data$weight,data=ind3data)

indo3data<-subset(wave3data,wave3data$country=="Indonesia")
indo3weight= svydesign(id=~1, weights=indo3data$weight,data=indo3data)

tan3data<-subset(wave3data,wave3data$country=="Tanzania")
t3weight= svydesign(id=~1, weights=tan3data$weight,data=tan3data)

u3data<-subset(wave3data,wave3data$country=="Uganda")
u3weight= svydesign(id=~1, weights=u3data$weight,data=u3data)

b3data<-subset(wave3data,wave3data$country=="Bangladesh")
b3weight= svydesign(id=~1, weights=b3data$weight,data=b3data)

p3data<-subset(wave3data,wave3data$country=="Pakistan")
p3weight= svydesign(id=~1, weights=p3data$weight,data=p3data)

#subset DFS users
dfsadopt_data <- subset(wave3data,wave3data$dfs_adopt=="1")
dfs_weight= svydesign(id=~1, weights=dfsadopt_data$weight, data = dfsadopt_data)

#subset non-DFS users
nodfs_data <- subset(wave3data, wave3data$dfs_adopt=="0")
nodfs_weight= svydesign(id=~1, weights = nodfs_data$weight, data = nodfs_data)
```

## Data Exploration

With the subsets of data, we can generate weighted proportions of our variables for each country in wave 3:
```{r wave 3 proportions}
#WAVE 3
prop_ed3=svyby(~ed_level, ~country, w3weight, svymean, na.rm = TRUE)
prop_emp3=svyby(~employed, ~country, w3weight, svymean, na.rm = TRUE)
prop_bank3=svyby(~bank_own, ~country, w3weight, svymean, na.rm = TRUE)
prop_marr3=svyby(~married, ~country, w3weight, svymean, na.rm = TRUE)
prop_lit3=svyby(~literate, ~country, w3weight, svymean, na.rm = TRUE)
prop_num3=svyby(~numerate, ~country, w3weight, svymean, na.rm = TRUE)
prop_sim3=svyby(~sim_ownoraccess, ~country, w3weight, svymean, na.rm = TRUE)
prop_marr3=svyby(~married, ~country, w3weight, svymean, na.rm = TRUE) 
prop_fem3=svyby(~female, ~country, w3weight, svymean, na.rm = TRUE)

prop_ed3 # example
```

Let's test for differences by education level:
```{r t test ed}
# education differences, adoption
tablee <- svytable(~ed_level+country, w3weight)
tablee
svychisq(~ed_level+country, w3weight)
```

Now testing for differences in DFS adoption rates by country:
```{r t test dfs}
# test for differences
table <- svytable(~dfs_adopt+country, w3weight) # cross tab 
summary(table, statistic="Chisq") # differences present by country
```

Differences for gender?
```{r t test gender}
# gender differences, adoption
tablef <- svytable(~dfs_adopt+female+country, w3weight)
tablef
svychisq(~dfs_adopt+female, w3weight) # over all countries, differences
```

Differences for ppi (poverty index) score?
```{r t test ppi}
# ppi differences, adoption
tableppi <- svytable(~dfs_adopt+country, w3weight)
t.test(wave3data$ppi_score~wave3data$dfs_adopt,var.equal = T)
```

## Plot education levels by country, DFS Users

Plotting education levels for DFS Users:
```{r ed plot}
proped_dfs=svyby(~ed_level, ~country, dfs_weight, svymean, na.rm = TRUE)

# plot details:
legendText=c("No Formal Education","Primary Education","Secondary Education and Above")
whereLegend="top"
shrinkLegend=0.8
showBorders=FALSE
groupColors=c("lightblue", "darkblue", "gray")

barplot(proped_dfs, ylim=c(0,1), border=showBorders, col=c("lightblue","darkblue","gray"), main="DFS User Education Levels by Country, Wave 3", bty="n", cex.main=3) # here it's proportions
# here comes the legend
legend(x=whereLegend, bty="n",
       legend = legendText, 
       fill = groupColors,
       cex=shrinkLegend)
```

## Plot histogram distributions for age

Plotting age distributions by country
```{r age plot}
# all together now
par(mfrow=c(2,4))
svyhist(~age, k3weight, main = "Kenya", col="blue", prob=FALSE)
svyhist(~age, n3weight, main = "Nigeria", col="blue", prob=FALSE)
svyhist(~age, ind3weight, main = "India", col="blue", prob=FALSE)
svyhist(~age, t3weight, main = "Tanzania", col="blue", prob=FALSE)
svyhist(~age, u3weight, main = "Uganda", col="blue", prob=FALSE)
svyhist(~age, p3weight, main = "Pakistan", col="blue", prob=FALSE)
svyhist(~age, b3weight, main = "Bangladesh", col="blue", prob=FALSE)
```

## Plot Gender of DFS Users

Plotting the gender breakdown of DFS users:
```{r Gender plot}
table_f=svytable(~female+country, dfs_weight)
summary(table_f)
propf_dfs=svyby(~female, ~country, dfs_weight, svymean, na.rm = TRUE)
whereLegend="topleft"
legendText=c("Male","Female")
groupColors=c("gray", "darkorchid4")
shrinkLegend=1.7
showBorders=FALSE

barplot(propf_dfs, beside=TRUE, border=showBorders, ylim=c(0,1), col=c("gray", "darkorchid4"), main="Gender of DFS Users, Wave 3", cex.main=3) 
legend(x= whereLegend, legend= legendText, fill= groupColors,
       cex=shrinkLegend, bty = "n")
```


## Calculating average DFS adoption rates

```{r DFS adoption rates}
# putting means for DFS adoption into separate frame
#means, wave 1
means1 <- svyby(~dfs_adopt,~country,design=w1weight,svymean,na.rm=TRUE)
means1

#means, wave 2
means2 <- svyby(~dfs_adopt,~country,design=w2weight,svymean,na.rm=TRUE)
means2

#means, wave 3
means3 <- svyby(~dfs_adopt, ~country, design = w3weight, svymean, na.rm=TRUE)
means3
```

## Aggregating Means together

Preparing average DFS adoption rates data for mapping:
```{r Prep dfs adoption data}
## aggregate means data into data frame ##
names(means2)[names(means2)=="dfs_adopt1"]="dfs_adopt2"
means2=means2[,c("country","dfs_adopt2")]
names(means3)[names(means3)=="dfs_adopt1"]="dfs_adopt3"
means3=means3[,c("country","dfs_adopt3")]

dfs_adopt_means=means1[,c("country","dfs_adopt1")]
dfs_adopt_means=merge(dfs_adopt_means,means2, by.x="country",by.y="country")
dfs_adopt_means=merge(dfs_adopt_means,means3, by.x="country", by.y = "country")

#change numbers to percentages
dfs_adopt_means$dfs_adopt1=dfs_adopt_means$dfs_adopt1*100
dfs_adopt_means$dfs_adopt2=dfs_adopt_means$dfs_adopt2*100
dfs_adopt_means$dfs_adopt3=dfs_adopt_means$dfs_adopt3*100

head(dfs_adopt_means)
```

## Mapping

Import map and prepare to plot data:
```{r get map}
# Get Map
library(maptools)
folder="Data/TM_WORLD_BORDERS-0.3/"
fileName="TM_WORLD_BORDERS-0.3.shp"
fileSHP=file.path(folder,fileName) 
globalmap =readShapeSpatial(fileSHP) 
head(globalmap@data)
plot(globalmap)
```

Merge map data and DFS adoption rates:
```{r map merge}
str(globalmap$NAME)
str(dfs_adopt_means)
globalmap =merge(globalmap,dfs_adopt_means,by.x='NAME',by.y='country',all.x=T)
```

Subset map for the regions with FII countries:
```{r subset map}
table(globalmap@data$REGION)
globalmap@data[,c('NAME','REGION')]

subMap=globalmap[globalmap@data$REGION%in%c('2','142'),]
```

Map DFS adoption rates by country, Wave 1:

```{r map1}
# make sure packages are installed
library(RColorBrewer)
library(classInt)

# wave 1
varToPLot=subMap@data$dfs_adopt1
numberOfClasses = 3
colorForScale='Greens'
title='DFS Adoption Rates by Country, Wave 1'

#plotting, wave 1
colors <- brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, style = "equal",dataPrecision=2)
colorPallette <- findColours(intervals, colors)
plot(subMap, col = colorPallette,main=title)
legend(-65,10, legend = names(attr(colorPallette, "table")), y.intersp=0.5, x.intersp=0.5, fill = attr(colorPallette, "palette"), cex = 0.75, bty = "n")

```

Wave 2:
```{r map2}
# wave 2
varToPLot=subMap@data$dfs_adopt2
numberOfClasses = 3
colorForScale='Greens'
title='DFS Adoption Rates by Country, Wave 2'

#plotting, wave 2
colors <- brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, style = "equal",dataPrecision=2)
colorPallette <- findColours(intervals, colors)
plot(subMap, col = colorPallette,main=title)
legend(-65,10, legend = names(attr(colorPallette, "table")), y.intersp=0.5, x.intersp=0.5, fill = attr(colorPallette, "palette"), cex = 0.75, bty = "n")

```

Wave 3:
```{r map3}
# wave 3
varToPLot=subMap@data$dfs_adopt3
numberOfClasses = 3
colorForScale='Greens'
title='DFS Adoption Rates by Country, Wave 3'

#plotting, wave 3
colors <- brewer.pal(numberOfClasses, colorForScale)
intervals <- classIntervals(varToPLot, numberOfClasses, style = "equal",dataPrecision=2)
colorPallette <- findColours(intervals, colors)
plot(subMap, col = colorPallette,main=title)
legend(-65,10, legend = names(attr(colorPallette, "table")), y.intersp=0.5, x.intersp=0.5, fill = attr(colorPallette, "palette"), cex = 0.75, bty = "n")
```

## Charts for Brochure

Adding a chart of unbanked populations to add detail to the brochure:
```{r unbanked plot}
banked=svyby(~bank_own,~country,design=w3weight,svymean, na.rm=TRUE)
legendText=c("Unbanked","Banked")
whereLegend="topleft"
shrinkLegend=1
groupColors=c("firebrick", "gray")
showBorders=FALSE

barplot(banked, beside=TRUE, ylim=c(0,1), border=showBorders, col=c("firebrick", "gray"), main="Chart 1: Unbanked vs. Banked Population, Wave 3", cex.main=2, cex.lab=3)
legend(x=whereLegend,
       legend = legendText, 
       fill = groupColors,
       cex=shrinkLegend, bty = "n")
```

## Plot PPI Scores

Plot PPI (Poverty Index) Score Distribution for DFS Users and Non-DFS Users: 

```{r boxplot}
# boxplot of dfs users
svyby(~ppi_score,~country,design=dfs_weight,svymean,na.rm=TRUE)
bp=svyboxplot(ppi_score~country, dfs_weight, all.outliers = FALSE, main = "Chart 2: PPI Poverty Scores, DFS Users, Wave 3", cex.main=2, col="darkolivegreen3")
# non-users
svyby(~ppi_score, ~country, design = nodfs_weight, svymean, na.rm=TRUE)
bp=svyboxplot(ppi_score~country, nodfs_weight, all.outliers = FALSE, main = "Chart 3: PPI Poverty Scores, Non-DFS Users, Wave 3", cex.main=2, col = "darkolivegreen")
```

